Neural networks for solving regression and classification problems

### 2022-11-17 doc/paper.tex

By introducing mini-batch gradient descent, we can observe how decreasing the batch size improves our performance further (Fig.~\ref{batch_constant_v_momentum}). Mini-batch gradient descent clearly yields a lower MSE for the test data than full batch gradient descent for the same amount of epochs. The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent. The test data MSE reached $0.044$, an impressive improvement over constant's previous result of $0.198$ MSE with full batch gradient descent. Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that these results will have some bias towards the batch size used when searching for the optimal hyperparameters. However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant. A full comparison of MSE over batch sizes for all the different schedulers can be viewed in Figure ~\ref{sgdbatch} in the appendix, but all schedulers yield increasingly better results as the batch size decreases.

\begin{figure}[H]
\centerline{\includegraphics[width=6in]{figure/SGD_batch_momentum_constant.png}}
\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent. A batch size of 320 is full batch GD, whilst a batch size of 1 is fully stochastic GD.}
\label{batch_constant_v_momentum}
\end{figure}

@@

@@

% \begin{figure}[H]
% \centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
% \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
% \label{adam_v_adagrad}
% \end{figure}

@@

>>>>

By introducing mini-batch gradient descent, we can observe how decreasing the batch size improves our performance further (Fig.~\ref{batch_constant_v_momentum}). Mini-batch gradient descent clearly yields a lower MSE for the test data than full batch gradient descent for the same amount of epochs. The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent. The test data MSE reached $0.044$, an impressive improvement over constant's previous result of $0.198$ MSE with full batch gradient descent. Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that these results will have some bias towards the batch size used when searching for the optimal hyperparameters. However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant. A full comparison of MSE over batch sizes for all the different schedulers can be viewed in Figure ~\ref{sgdbatch} in the appendix, where we can see that all schedulers yield increasingly better results as the batch size decreases.

\begin{figure}[H]
\centerline{\includegraphics[width=6in]{figure/new_SGD_batch_momentum_constant.png}}
\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent. A batch size of 320 is full batch GD, whilst a batch size of 1 is fully stochastic GD.}
\label{batch_constant_v_momentum}
\end{figure}

@@

Based on these findings, we decided to push Adagrad and Adam further, allowing them to run for much longer. After 5000 epochs, we reached an MSE of $0.012$ for Adam and $0.011$ for Adagrad. The plot for the MSE over epochs can be found in the appendix (Fig.~\ref{adam_v_adagrad}).

@@

@@

\begin{figure}[H]
\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
\caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
\label{adam_v_adagrad}
\end{figure}
